Hunting experience and prey variability jointly shape individual foraging specialisation in a predator-prey videogame

### 2023-01-10 manuscript/Chapter2-manuscript.Rmd
update styles, ms, and ms rmd

In this study, we analyze individual behavioural data from players in *Dead by Daylight* (*DBD*) to test how hunting expertise and prey behaviour shape predator foraging specialization. *DBD* is an online videogame simulating a predator-prey interaction, where one predator player hunts four prey players in different virtual environments. The data grants a high degree of precision on the behavioural interaction as both the predator and the four prey are measured simultaneously in each trial. First, we investigate how predators developed their individual hunting expertise. We hypothesize that predators should differ in the development of their expertise because they encounter varying levels of difficulty with the prey that they pursue. Second, we test the hypothesis that individual variation in foraging behaviour will change with experience. If experience reduces the costs of switching between hunting tactics, we predict that individual predators should become more flexible with time. Alternatively, if experience enables the refinement of the hunting tactics, then individuals may instead specialize. Third, we evaluate how predator foraging specialization interacts with variation in prey behaviour. Whether predators specialize or not with experience may depend on the behaviour of their prey, for instance, as it may be harder to specialize when encounters are less predictable. Lastly, if individuals differ in their degree of foraging specialization, then we expect specialist hunters to fare better when prey variability is lower, while flexible hunters should fare better when prey variability is higher.

@@

Our results confirm that individual predators displayed important differences in the development of their expertise, as the best model was the one with group-level smoothers exclusively (Table I). For instance, some predators steadily increased their hunting success with experience, while others maintained either a high or low success across their experience (Figure 1A). Moreover, some individuals show a decrease in hunting success with experience. Another pattern that emerged was that some individuals optimized their success at an experience level that was below what we considered as "advanced" (i.e. >300 matches played).

@@

Paragraphe qui parle des différences entre figure 1 A et B, soit, lorsqu'on contrôle pour l'effet des proies.

>>>>

In this study, we analyze individual behavioural data from players in *Dead by Daylight* (*DBD*) to test how hunting experience and prey behaviour shape predator foraging specialization. *DBD* is an online videogame simulating a predator-prey interaction, where one predator player hunts four prey players in different virtual environments. The data grants a high degree of precision on the behavioural interaction as both the predator and the four prey are measured simultaneously in each trial. First, we investigate how predators developed their individual hunting expertise. We hypothesize that predators should differ in the development of their expertise because they encounter varying levels of difficulty with the prey that they pursue. Second, we test the hypothesis that individual variation in foraging behaviour will change with experience. If experience reduces the costs of switching between hunting tactics, we predict that individual predators should become more flexible with time. Alternatively, if experience enables the refinement of the hunting tactics, then individuals may instead specialize. Third, we evaluate how predator foraging specialization interacts with variation in prey behaviour. Whether predators specialize or not with experience may depend on the behaviour of their prey, for instance, as it may be harder to specialize when encounters are less predictable. Lastly, if individuals differ in their degree of foraging specialization, then we expect specialist hunters to fare better when prey variability is lower, while flexible hunters should fare better when prey variability is higher.

@@

Our results confirm that individual predators displayed important differences in the development of their expertise, as the best model was the one with group-level smoothers exclusively (Table I). For instance, with experience, some predators steadily increased their hunting success while others displayed a decrease in prey capture (Figure 1A). Otherwise, there were players that maintained either a high or low success across their experience (Figure 1A). Another pattern that emerged was that some individuals optimized their success at an experience level that was below what we considered as "advanced" (i.e. >300 matches played).

@@

We predicted that the relationship between hunting success and experience would differ among predators because they should encounter different prey. However, after controlling for prey speed, we did not find evidence confirming our hypothesis, such that differences among predators in the relationship between hunting success and experience remained mostly the same (Figure 1B).

```{r fig1, out.width = "100%", fig.cap = "**Figure 1.** Among individual differences in the development of hunting expertise. The predators' hunting success (i.e. the probability of capturing the four prey) is on the y axis, and the predators' cumulative experience (i.e. the number of matches played prior to each observation) is on the x axis. Each fitted curve represents an individual predator. (A) A generalized additive mixed model where we do not control for prey speed (B) A generalized additive model where we control for prey speed."}

knitr::include_graphics(here::here("outputs", "05_outputs_figures", "05_figure1.png"))
